---
title: "setting-bins"
author: "devin"
date: "November 15th, 2015"
output: pdf_document
---

Setting bins in R, while not particulary difficult, is usually reasonably standardized 
and therefore can be easily automatted for many situations.

```{r}
library(PKPDdatasets)
library(PKPDmisc)
library(dplyr)
library(knitr)
```

```{r}
data <- capitalize_names(sd_oral_richpk)
sid_data <- data %>% filter(!duplicated(ID))
```

Given a traditional PK dataset (data) and one filtered for one observation per individual (sid_data) that looks like:

```{r}
head(data) %>% kable(digits =2)
```

\newpage

```{r}
head(sid_data) %>% kable(digits =2)
```

Here are a couple situations in which bins are appropriate


1. given a breakpoint weight of 70 kg, determine how flag any individuals are above the breakpoint weight.

Historically, this has been accomplished using ifelse statments like so
```{r}
binned_data <- sid_data %>% 
  mutate(WTBIN = ifelse(WEIGHT > 70, 1, 0))
head(binned_data) %>% kable(digits =2)
```

This can be similarly accomplished with binning using the `set_bins` function from `PKPDmisc`


```{r}
binned_data <- binned_data %>% 
  mutate(WTBIN2 = set_bins(WEIGHT, breaks = 70))

head(binned_data)%>% kable(digits =2)
```

The 'problem' with the ifelse statement is it does not scale well under more complex circumstances.

For example, given a more complex set of criteria such as weight less than 50, between 50 and 70, and 70 and above,
the above example must be expanded to:

```{r}
binned_data <- sid_data %>% 
  mutate(WTBIN = ifelse(WEIGHT < 50, 0, 
                        ifelse(WEIGHT >= 50 & WEIGHT < 70, 1,
                               ifelse(WEIGHT >= 70, 2, NA))))
head(binned_data)%>% kable(digits =2)
```

However, with the `set_bins` function and the `breaks` argument this is simple:

```{r}
binned_data <- binned_data %>% 
  mutate(WTBIN2 = set_bins(WEIGHT, breaks = c(50, 70)))
head(binned_data) %>% kable(digits =2)
```

\newpage

You can even find additional information about the ranges of each bin by setting `quiet=FALSE`

```{r}
 binned_data %>% 
  mutate(WTBIN2 = set_bins(WEIGHT, 
                           breaks = c(50, 70), 
                           quiet=F)) %>% 
  head %>% kable(digits =2)
```

As you can see by default, the first and last bin are automatically set to a range of -Inf and Inf for
the min and max values. This can be controlled using `lower_boundary` and `upper_boundary` such that
the lower and upper bounds can either be controlled, and any values exceeding those bounds will be returned as `NA`. This might be useful if you are programmatically calculating the break points and already will have a lower bound specified. 

For example by manually specifying a lower bound of 0:

```{r}
 binned_data %>% 
  mutate(WTBIN2 = set_bins(WEIGHT, 
                           breaks = c(0, 50, 70), 
                           quiet=F)) %>% 
  head %>% kable(digits =2)
```

It adds an unnecessary -Inf to 0 bin. This can be turned off via `lower_bound=NULL`:

```{r}
 binned_data %>% 
  mutate(WTBIN2 = set_bins(WEIGHT, breaks = c(0, 50, 70), 
                           lower_bound=NULL,  
                           quiet=F)) %>% 
  head %>% kable(digits =2)
```

## Getting away from `quantile` and `cut`

A common way of stratifying is to use quantile and the cut function to establish bins. The cut function, however,
has poor defaults, such as not including the lowest value, and returns a string column specifying the range
in each cut that must be further coerced to arrive at a numeric bin.

```{r}
sid_data %>% 
  mutate(AGECUTS = cut(AGE, 
                       breaks = quantile(AGE), 
                       include.lowest=T),
                    AGEBINS = as.numeric(AGECUTS)) %>% 
  head %>% kable(digits =2)


```

This is again simplified, especially as, by default, the breaks argument calculates the quantiles

```{r}
sid_data %>% mutate(AGEBINS = set_bins(AGE)) %>%
  head %>% kable(digits =2)
```


For more fine-tuned levels, we can specify additional breaks

```{r}
sid_data %>% 
  mutate(AGEBINS = set_bins(AGE, 
                            breaks = quantile(AGE, 
                                              seq(0, 1, length.out = 10)
                                              ),
                           quiet = F)
         ) %>% 
  head %>% 
  kable(digits =2)

```



## BETWEEN

A special case is when there is desire to test whether a value falls in some inclusive range, such
as a specified therapeutic window. `set_bins` can also handle this situation with the `between` argument.

In this case, a range can be specified, and all values inside the range (inclusive) will be assigned to bin 1, 
with all values outside in either bin 0 (below range) or bin 2 (above range).

For example, to understand the non-zero concentration measurements given a therapeutic window of 20-100, the following
can be done:

```{r}
tw_data <- data %>% 
  filter(CONC > 0) %>% 
  mutate(TW = set_bins(CONC, between=c(20, 100)))

head(tw_data) %>% kable(digits =2)
```

This can easily be further examined:

```{r}
tw_data %>% group_by(TW) %>% 
  summarize(n= n()) %>% 
  kable(digits =2)

tw_data %>% filter(TW ==2) %>% kable(digits =2)
```

\newpage

```{r}
devtools::session_info()
```

