---
title: "broom"
author: "Devin"
date: "March 29, 2015"
output: html_document
---

```{r}
library(PKPDmisc)
library(dplyr)
library(broom)
```
```{r}
FUdata <- read_table("CS2.csv")
```

```{r}
fudata <- FUdata %>% mutate(AGE1 = AGE/50,
                  BSA1 = BSA/1.7,
                  DOSE1 = DOSE/1000)
```

```{r}
mb_clfit <- lm(log(FU_CL) ~ AGE1+BSA1+DOSE1+factor(SEX)+factor(MTX),data = fudata)
```

# default information

Is not particularly programmatically designed, making it difficult to access elements, and you must visually search through lines

```{r}
mb_clfit
```

```{r}
summary(mb_clfit)
```

# With broom

Some key functions broom offers:

### Tidy

```{r}
tidy(mb_clfit)
```

can even turn it into a nice table with knitr's `kable` function and using `results='asis'` in the rmarkdown chunk

```{r, results='asis'}
knitr::kable(tidy(mb_clfit))
```


### glance

access the key statistics quickly

```{r}
glance(mb_clfit)
```

can pull them out by name easily
```{r}
glance(mb_clfit)$r.squared
```


### augment

Gives you the fitted values, and makes it very easy to turn them into a plot

```{r}
aug <- augment(mb_clfit)
head(aug)
```

```{r}
library(ggplot2)
ggplot(data = aug, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  base_theme() +
  geom_hline(yintercept = 0, size = 1.25) +
  labs(list(x = "Predicted log(CL)", y = "Residuals"))
```

