---
title: "cut_function"
author: "Devin"
date: "Tuesday, June 17, 2014"
output: html_document
---

Default of cut breaks to be closed for lowest and open for highest, this will cause the individual with the lowest value to not be assigned a level properly
```{r}
wt_quartiles <- quantile(Theoph$Wt)
wt_quartiles
Theoph$WTBINS <- cut(Theoph$Wt, breaks = wt_quartiles)

# will not bin person with lowest weight
head(subset(Theoph, Wt == wt_quartiles[1]))

# heaviest person gets binned properly
head(subset(Theoph, Wt == wt_quartiles[length(wt_quartiles)]))
```

Use argument `include.lowest` to include lowest (or highest if `right = FALSE`)

```{r}
Theoph$WTBINS_ALL <- cut(Theoph$Wt, breaks = wt_quartiles, include.lowest = TRUE)

# WTBINS_ALL will bin person with lowest weight
head(subset(Theoph, Wt == wt_quartiles[1]))

# heaviest person still gets binned properly
head(subset(Theoph, Wt == wt_quartiles[length(wt_quartiles)]))

levels(Theoph$WTBINS_ALL)
```

Can extract the lower and upper bound
```{r}
# from cut documentation
lower <-  as.numeric( sub("\\((.+),.*", "\\1", levels(Theoph$WTBINS)))
upper <-  as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", levels(Theoph$WTBINS)))
lower
upper
```

however if `include.lowest = TRUE` the lower argument will have an NA as for the lower bound for the lowest level 
```{r}

lower <-  as.numeric( sub("\\((.+),.*", "\\1", levels(Theoph$WTBINS_ALL)))
upper <-  as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", levels(Theoph$WTBINS_ALL)))
lower
upper
```


TODO: Break down the regular expression


